---
title: "Lab Assignment 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Practise

Installing packages
```{r}
install.packages('rgdal')
install.packages('raster')	
install.packages("GISTools")
install.packages('rgeos')
install.packages('maptools')
install.packages('mapview')
```
Loading libraries
```{r}
library(raster)
library(rgdal)
library(GISTools)
library(rgeos)
library(maptools)
library(mapview)
```

Reading the shape file 
```{r}
pophu<-readOGR('.',layer='y2010_51_pophu')

```

ploting the shape file
```{r}
plot(pophu)
```

Plotting shap file with attributes
```{r}
choropleth(pophu,pophu$POP10)
```
```{r}
choropleth(pophu,pophu$HOUSING10)
```


Filtering
```{r}
pophu_sub <- pophu[pophu$POP10>100,]
```

Rasters
Reading raster
```{r}
nlcd<-raster('nlcd.tif')
nlcd
```
ploting
```{r}
plot(nlcd)
mapview(nlcd)
```
Obtain Raster Attributes
```{r}
nlcd
is.factor(nlcd)

```
clipping
```{r}
Montgomery <- countries[countries$NAME=='Montgomery',] 
nlcdsub <- crop(nlcd, extent(Montgomery))
plot(nlcdsub)
plot(Montgomery,add=TRUE)

```
Masking
```{r}
nlcdsub <- mask(nlcdsub, Montgomery)
plot(nlcdsub)
```
Use length() function to report total pixel numbers for a specific land cover type. 
```{r}
length(nlcdsub[nlcdsub==11])
```

masking and ploting as per criteria
```{r}
forest <- nlcdsub
forest[forest>=41&forest<=43]=1
forest[forest!=1]=0
plot(forest)
```






## Assignment 
### Question one
```{r}
countries<-readOGR('.',layer='counties')
```


### Question two:
2000
```{r}
sum(countries$POP2000)
mean(countries$POP2000)
```
2003
```{r}
sum(countries$POP2003)
mean(countries$POP2003)
```

### Question 3
```{r}
M_P<-countries[countries$NAME=="Montgomery"|countries$NAME=="Pulaski",]
plot(M_P)+polygonsLabel(M_P,M_P$NAME)
```

### Question 4

```{r}
nlcd<-raster('nlcd.tif')
Montgomery <- countries[countries$NAME=='Montgomery',] 
nlcdsub <- crop(nlcd, extent(Montgomery))
nlcdsub <- mask(nlcdsub, Montgomery)
plot(nlcdsub)
```

Deciduous forest (code=41)
```{r}
length(nlcdsub[nlcdsub==41])
length(nlcdsub[nlcdsub==41])*xres(nlcdsub)*yres(nlcdsub)/1e6
```

Evergreen forest (code=42)
```{r}
length(nlcdsub[nlcdsub==42])
length(nlcdsub[nlcdsub==42])*xres(nlcdsub)*yres(nlcdsub)/1e6
```

Mixed forest (code=43)
```{r}
length(nlcdsub[nlcdsub==43])
length(nlcdsub[nlcdsub==43])*xres(nlcdsub)*yres(nlcdsub)/1e6
```


### Question5

```{r}
urban=nlcdsub
urban[urban>=21&urban<=24]=1
urban[urban!=1]=0
plot(urban)
```





































